How accurate is CC4 for vertical excitation energies?

### 2021-05-01 Manuscript/CC4.tex
CC4 data

\section{Introduction}

The hierarchy of coupled cluster (CC) methods provide a route to systematically improvable energies and properties via the systematic increase of the maximum excitation degree of the cluster operator $\hT = \hT_1 + \hT_2 + \ldots + \hT_N$ (where $N$ is the number of electrons).
Without any truncation, full CC (FCC) is equivalent to full configuration interaction (FCI) and provides the exact energy of the system. 
However, it is not computationally viable and one usually resorts to truncated CC methods.
Popular choices are CC with singles and doubles (CCSD), CC with singles, doubles, and triples (CCSDT), CC with singles, doubles, triples, and quadruples (CCSDTQ), and 
CC with singles, doubles, triples, quadruples, and pentuples (CCSDTQP) with corresponding computational scalings of $\order*{M^{6}}$, $\order*{M^{8}}$,  $\order*{M^{10}}$, and  $\order*{M^{12}}$, respectively (where $M$ denotes the number of orbitals).
An alternative systematically-improvable family of methods is provided by the CC2, CC3, and CC4 models which have been introduced by the Aarhus group in the context of CC response theory.
These methods scale as $\order*{M^{5}}$, $\order*{M^{7}}$, and $\order*{M^{9}}$, respectively, and can be seen as a cheap approximation of CCSD, CCSDT, and CCSDTQ, by skipping the most expensive terms and avoiding the storage of the higher-excitation amplitudes.
A similar strategy has been applied to produce the CCSDT-3 and CCSDTQ-3 models based on arguments stemming from perturbation theory.
Of course, many more approximate CC methods have been developed and we refer the interested reader to specialized reviews for more details.

Coupled-cluster methods have been particularly successful for small- and medium-sized molecules in the field of thermodynamics and spectroscopy, thanks to the computations of accurate geometries, potential energy surfaces, couplings between electronic states, a vast panel of properties such as oscillator strengths as well as ground- and excited-state dipoles.
Although originally developed from ground-state energies and properties, CC has been successfully extended to excited states thanks to the equation-of-motion (EOM) and linear-response (LR) formalisms which are known to produce identical excitation energies but different properties.
In EOM-CC, one obtains vertical excitation energies via the diagonalization of the similarity-transformed Hamiltonian $\bH = e^{-\hT} \hH e^{\hT}$.
A more general procedure to compute excitation energies that can be straightforwardly applied to any truncated CC family as the ones mentioned above consists in diagonalizing the so-called CC Jacobian obtained via the differentiation of the CC amplitude equations.

@@

\section{Computational details}

Each method used in the present study is reported in Table \ref{tab:scaling} alongside its formal computational scaling and the electronic structure software used to compute the excitation energies.

\section{Results and discussion}

@@

\section{Conclusion}

@@

\begin{acknowledgements}
This work was performed using HPC resources from CALMIP (Toulouse) under allocation 2021-18005.
PFL has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (Grant agreement No.~863481).
\end{acknowledgements}

\section*{Data availability statement}
The data that support the findings of this study are openly available in Zenodo at \href{http://doi.org/XX.XXXX/zenodo.XXXXXXX}{http://doi.org/XX.XXXX/zenodo.XXXXXXX}.

\bibliography{CC4}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Coupled cluster (CC) theory provides a hierarchy of methods that provide systematically improvable energies and properties via the systematic increase of the maximum excitation degree of the cluster operator $\hT = \hT_1 + \hT_2 + \ldots + \hT_N$ (where $N$ is the number of electrons).
Without any truncation, full CC (FCC) is equivalent to full configuration interaction (FCI) and provides the exact energy of the system. 
However, it is not computationally viable and one usually resorts to truncated CC methods.
Popular choices are CC with singles and doubles (CCSD), CC with singles, doubles, and triples (CCSDT), CC with singles, doubles, triples, and quadruples (CCSDTQ), and 
CC with singles, doubles, triples, quadruples, and pentuples (CCSDTQP) with corresponding computational scalings of $\order*{M^{6}}$, $\order*{M^{8}}$,  $\order*{M^{10}}$, and  $\order*{M^{12}}$, respectively (where $M$ denotes the number of orbitals).
An alternative systematically-improvable family of methods is provided by the CC2, CC3, and CC4 models which have been introduced by the Aarhus group in the context of CC response theory.
These methods scale as $\order*{M^{5}}$, $\order*{M^{7}}$, and $\order*{M^{9}}$, respectively, and can be seen as a cheap approximation of CCSD, CCSDT, and CCSDTQ, by skipping the most expensive terms and avoiding the storage of the higher-excitation amplitudes.
A similar strategy has been applied to produce the CCSDT-3 and CCSDTQ-3 models based on arguments stemming from perturbation theory.
Of course, many more approximate CC methods have been developed and we refer the interested reader to specialized reviews for more details.

Coupled-cluster methods have been particularly successful for small- and medium-sized molecules in the field of thermodynamics, kinetics, and spectroscopy, thanks to the computations of accurate geometries, potential energy surfaces, couplings between electronic states, a vast panel of properties such as dipoles and higher moments, vibrational frequencies, NMR chemical shifts, magnetizabilities, polarizabilities, etc.
Although originally developed from ground-state energies and properties, CC has been successfully extended to excited states thanks to the equation-of-motion (EOM) and linear-response (LR) formalisms which are known to produce identical excitation energies but different properties.
In EOM-CC, one obtains vertical excitation energies via the diagonalization of the similarity-transformed Hamiltonian $\bH = e^{-\hT} \hH e^{\hT}$.
A more general procedure to compute excitation energies that can be straightforwardly applied to any truncated CC family as the ones mentioned above consists in diagonalizing the so-called CC Jacobian obtained via the differentiation of the CC amplitude equations.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Computational details}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Each method used in the present study is reported in Table \ref{tab:scaling} alongside its formal computational scaling and the electronic structure software used to compute the excitation energies.
In a nutshell, we have used by default the {\CFOUR} program to compute the CC energies at the notable exception of CCSDTQP where we have employed {\MRCC} and its automated implementation of high-order CC methods based on the strings of spin-orbital indices.
The FCI estimates were obtained with the CIPSI algorithm implemented in {\QP}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgements}
This work was performed using HPC resources from CALMIP (Toulouse) under allocation 2021-18005.
PFL has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (Grant agreement No.~863481).
\end{acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Data availability statement}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The data that support the findings of this study are openly available in Zenodo at \href{http://doi.org/XX.XXXX/zenodo.XXXXXXX}{http://doi.org/XX.XXXX/zenodo.XXXXXXX}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{CC4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
